The piece is scored for stage actor, alto flute, oboe, two guitars, violin and
cello.

\textbf{Accidentals.} Accidentals govern only one note. This is true even for
successive noteheads at the same staff position. The sequence of G$\sharp$4
followed by G4 (without accidental) is to be understood as G$\sharp$4 followed
by G$\natural$4.

\textbf{Appoggiaturas.} Play runs of small-note appoggiaturas as fast as
possible starting directly on the beat; land immediately on the full-size note
shown below and sustain to the end of the duration indicated.

\textbf{Glissandi.} Although the start-pitches of glissandi should be clearly
articulated, neither the stop-pitch nor any intermediate pitches need to be
rearticulated. This is as true for single-segment portamenti in the alto flute
as for the elaborate multistage glissandi in the violin and cello. In other
words, all nonfirst pitches in glissandi may be thought of as being small
parenthesized notes (whether they are in fact notated this way or not).
Quarter-note stems are included to clarify the rhythm of some longer glissandi;
these stems are rhythmic indicators only and do not change the sound of the
glissandi produced. Play all glissandi as smooth and continuous, rather than
angular or disjointed.

\textbf{Flat glissandi.} Flat glissandi are sometimes used as a typographical
variant of ties.

\textbf{``Terminate abruptly.''} The symbol of a capital T turned on its side
is used to mean that the sound so-marked is to terminated abruptly (ripped or
``gerisst''), rather than being allowed to fade away of its own accord. To
accomplish this, the flute should bring the tongue to the back of the teeth (or
alveolar ridge of the gums) to stop the air; the oboe should touch the tongue
to the reed to stop the air; and the violin and cello use the bow to choke the
sound at the end of its duration, termating such notes stop-on-string. The
symbol is written for the guitars only over string-scraping gestures; terminate
these in similar fashion to the violin and cello, using the scraper to suddenly
choke the sound, marking the end of the note stop-on-string. In all cases, the
sound ideal is of a decidedly `square' or `rectangular' sound that treats the
termination of the sound a distinct event.

\textbf{Winds.} Trills without secondary pitches are bisbigliandi, or color
trills. Pick a secondary fingering that nearly matches the pitch of the primary
note. If no color fingering is available, play the secondary pitch a minor
second higher than the primary pitch. The alto flute sounds a perfect fourth
lower than written. Whistle tones are shown as runs of small note-heads written
with neither stems nor dynamics; finger these over the pitch-class of the first
note in such passages. Play passages marked ``covered'' (or ``cov.'') by
covering the opening of the flute with the lips, sounding a perfect fourth plus
a major seventh lower than written; such passages sometimes (though not always)
are notated as dyads to show the sounding pitch. Airtone is marked as such,
sometimes abbreviated ``air.''

\textbf{Guitars.} No scordatura; the guitars sound an octave lower than
written. In general, the sound ideal for the guitars is as resonant and
luminous as possible. Although many lone notes are equipped with laissez-vibrer
ties, essentially all passages in the guitars can be played with glowing
resonance. This is especially true prior to fermatas and other silences, where
after-echo of the guitars is especially to be encouraged. \textbf{Doits and
falls.} Doits and falls are usually written at an interval of a fourth; follow
the up- or down-direction of doits and falls, but stretch or compress the
distance of these figures \textit{ad lib}. \textbf{Lateral scraping up the
string.} Play passages written on a 1-line staff by scraping laterally up the
wire wrapping of the low E-string with a credit card, plastic pick, small coin
or other item. The goal is a rhythmically precise burst of ``ribbed'' or
``serrated'' or ``guiro-like'' white noise. (These passages frequently
articulate a type of rhythmic canon between the two guitars.) Follow the up-bow
and down-down indications given. \textbf{Metal-screw ``rasgueado.''} At the end
of each of the four ``Stills'' in the piece a note is marked ``with screw.''
The indication here means to finger the pitch indicated while bowing the low E
string with a long, lightweight metal machinist's screw, available at most
hardware stores. Make sure the screw is at least 8 -- 12 inches long, which
seems to help with the movement of the screw over the wraping of the string.
When bowing, draw the screw across the string at a spot only a fret or two away
from the finger of the left hand: trial and error will reveal at string contact
point that elicits a cascade of glowing metallic ``flecks'' in a dancing
spectrum of sound rooted at the pitch fingered by the left hand. Jürgen Ruck
played a first version of this sound in my piece \textit{L'archipel du corps}
in 2011. Nico Couck developed the version of the sound called for here in
\textit{Spiel der Dornen}, premiered at Darmstadt in 2016.
